### 2019-01-22 paper/data_template.tex
changed up some of the output to support mana's queries

\multirow{2}{*}{Ranking System}&\multicolumn{4}{c||}{Case Rank - Median (Mean)}&\multicolumn{4}{c}{Percentage in Top X Variants - X=(1, 10, 20)} \\ \cline{2-9}
&All (n={{ TEST_COUNTS.OVERALL }})&
VUS (n={{ TEST_COUNTS.VARIANT_OF_UNCERTAIN_SIGNIFICANCE }})&
LP (n={{ TEST_COUNTS.LIKELY_PATHOGENIC }})&Path. (n={{ TEST_COUNTS.PATHOGENIC }})&
All (n={{ TEST_COUNTS.OVERALL }})&
VUS (n={{ TEST_COUNTS.VARIANT_OF_UNCERTAIN_SIGNIFICANCE }})&
LP (n={{ TEST_COUNTS.LIKELY_PATHOGENIC }})&Path. (n={{ TEST_COUNTS.PATHOGENIC }}) \\ \hhline{=|=|=|=|=||=|=|=|=}
{% for cs_label in CS_LABELS %}
    {{ cs_label }}&
    {% for LEVEL in LEVELS %}
        {{ '%0.1f (%0.2f)' % (np.median(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]), np.mean(COMPARISON[cs_label].TEST_RANKINGS[LEVEL])) }}&
    {% endfor %}
    {% for LEVEL in LEVELS %}
        {{ '%d, %d, %d' % 
            (100.0*np.sum(np.array(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]) <= 1) / len(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]),
            100.0*np.sum(np.array(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]) <= 10) / len(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]),
            100.0*np.sum(np.array(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]) <= 20) / len(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]))
        }}
        {% if LEVEL != 'PATHOGENIC' %}
            &
        {% endif %}
    {% endfor %}
    \\
{% endfor %}
\hhline{=|=|=|=|=||=|=|=|=}

Exomiser(best)&
{% for LEVEL in LEVELS %}
    {{ '%0.1f (%0.2f)' % (np.median(EXOMISER_BEST.TEST_RANKINGS[LEVEL]), np.mean(EXOMISER_BEST.TEST_RANKINGS[LEVEL])) }}&
{% endfor %}
{% for LEVEL in LEVELS %}
    {{ '%d, %d, %d' % 
        (100.0*np.sum(np.array(EXOMISER_BEST.TEST_RANKINGS[LEVEL]) <= 1) / len(EXOMISER_BEST.TEST_RANKINGS[LEVEL]),
        100.0*np.sum(np.array(EXOMISER_BEST.TEST_RANKINGS[LEVEL]) <= 10) / len(EXOMISER_BEST.TEST_RANKINGS[LEVEL]),
        100.0*np.sum(np.array(EXOMISER_BEST.TEST_RANKINGS[LEVEL]) <= 20) / len(EXOMISER_BEST.TEST_RANKINGS[LEVEL]))
    }}
    {% if LEVEL != 'PATHOGENIC' %}
        &
    {% endif %}
{% endfor %}
\\

Exomiser(avg)&
{% for LEVEL in LEVELS %}
    {{ '%0.1f (%0.2f)' % (np.median(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL]), np.mean(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL])) }}&
{% endfor %}
{% for LEVEL in LEVELS %}
    {{ '%d, %d, %d' % 
        (100.0*np.sum(np.array(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL]) <= 1) / len(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL]),
        100.0*np.sum(np.array(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL]) <= 10) / len(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL]),
        100.0*np.sum(np.array(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL]) <= 20) / len(EXOMISER_AVERAGE.TEST_RANKINGS[LEVEL]))
    }}
    {% if LEVEL != 'PATHOGENIC' %}
        &
    {% endif %}
{% endfor %}

\\ \hhline{=|=|=|=|=||=|=|=|=}

{% for clf_label in CLF_LABELS %}
    {{ clf_label }}&
    {% for LEVEL in LEVELS %}
        {{ '%0.1f (%0.2f)' % (np.median(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]), np.mean(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL])) }}&
    {% endfor %}
    {% for LEVEL in LEVELS %}
        {{ '%d, %d, %d' % 
            (100.0*np.sum(np.array(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]) <= 1) / len(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]),
            100.0*np.sum(np.array(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]) <= 10) / len(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]),
            100.0*np.sum(np.array(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]) <= 20) / len(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]))
        }}
        {% if LEVEL != 'PATHOGENIC' %}
            &
        {% endif %}
    {% endfor %}
    \\
{% endfor %}