### 2019-03-19 paper/supplements/hyperparameter_template.tex

{% for clf_label in CLF_LABELS %}
    {{ '\\multirow{%d}{*}{\\parbox{1.0\\linewidth}{%s}}' % (len(CLASSIFIERS[clf_label]['BEST_PARAMS']), clf_label.replace('(', ' (')) }} 
    {% for k in CLASSIFIERS[clf_label]['BEST_PARAMS'] %}
        &{{ k.replace('_', '\\_') }}&{{ CLASSIFIERS[clf_label]['HYPERPARAMETER_SPACE'][k] }}&{{ CLASSIFIERS[clf_label]['BEST_PARAMS'][k] }} \\ \cline{2-4}
    {% endfor %}
    \hline
{% endfor %}

%resultsDict['CLASSIFIERS'][clf_label]['BEST_PARAMS'] = clf.best_params_
%resultsDict['CLASSIFIERS'][clf_label]['HYPERPARAMETER_SPACE'] = hyperparam