### 2019-04-10 paper/data_template.tex
fixed templating with pathogenic only

\multirow{2}{*}{Ranking System}&\multicolumn{4}{c||}{Case Rank - Median (Mean)}&\multicolumn{4}{c}{Percentage in Top X Variants - X=(1, 10, 20)} \\ \cline{2-9}
&All (n={{ TEST_COUNTS.OVERALL }})&
VUS (n={{ TEST_COUNTS.get('VARIANT_OF_UNCERTAIN_SIGNIFICANCE', 0) }})&
LP (n={{ TEST_COUNTS.LIKELY_PATHOGENIC }})&Path. (n={{ TEST_COUNTS.PATHOGENIC }})&
All (n={{ TEST_COUNTS.OVERALL }})&
VUS (n={{ TEST_COUNTS.get('VARIANT_OF_UNCERTAIN_SIGNIFICANCE', 0) }})&
LP (n={{ TEST_COUNTS.LIKELY_PATHOGENIC }})&Path. (n={{ TEST_COUNTS.PATHOGENIC }}) \\ \hhline{=|=|=|=|=||=|=|=|=}
{% for cs_label in CS_LABELS %}
    {{ cs_label }}&
    {% for LEVEL in LEVELS %}
        {% if TEST_COUNTS.get(LEVEL, 0) == 0 %}
            -- (--)&
        {% else %}
            {{ '%0.1f (%0.2f)' % (np.median(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]), np.mean(COMPARISON[cs_label].TEST_RANKINGS[LEVEL])) }}&
        {% endif %}
    {% endfor %}
    {% for LEVEL in LEVELS %}
        {% if TEST_COUNTS.get(LEVEL, 0) == 0 %}
            --
        {% else %}
            {{ '%d, %d, %d' % 
                (100.0*np.sum(np.array(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]) <= 1) / len(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]),
                100.0*np.sum(np.array(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]) <= 10) / len(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]),
                100.0*np.sum(np.array(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]) <= 20) / len(COMPARISON[cs_label].TEST_RANKINGS[LEVEL]))
            }}
        {% endif %}
        {% if LEVEL != 'PATHOGENIC' %}
            &
        {% endif %}
    {% endfor %}
    \\
{% endfor %}
\hhline{=|=|=|=|=||=|=|=|=}

{% for ext_label in EXT_LABELS %}
    {{ ext_label.replace('_', '\\_') }}&
    {% for LEVEL in LEVELS %}
        {% if TEST_COUNTS.get(LEVEL, 0) == 0 %}
        -- (--)&
        {% else %}
            {{ '%0.1f (%0.2f)' % (np.median(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL]), np.mean(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL])) }}&
        {% endif %}
    {% endfor %}
    {% for LEVEL in LEVELS %}
        {% if TEST_COUNTS.get(LEVEL, 0) == 0 %}
            --
        {% else %}
            {{ '%d, %d, %d' % 
                (100.0*np.sum(np.array(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL]) <= 1) / len(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL]),
                100.0*np.sum(np.array(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL]) <= 10) / len(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL]),
                100.0*np.sum(np.array(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL]) <= 20) / len(EXTERNAL[ext_label].TEST_RANKINGS[LEVEL]))
            }}
        {% endif %}
        {% if LEVEL != 'PATHOGENIC' %}
            &
        {% endif %}
    {% endfor %}
    \\
{% endfor %}
\hhline{=|=|=|=|=||=|=|=|=}

{% for clf_label in CLF_LABELS %}
    {{ clf_label }}&
    {% for LEVEL in LEVELS %}
        {% if TEST_COUNTS.get(LEVEL, 0) == 0 %}
        -- (--)&
        {% else %}
            {{ '%0.1f (%0.2f)' % (np.median(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]), np.mean(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL])) }}&
        {% endif %}
    {% endfor %}
    {% for LEVEL in LEVELS %}
        {% if TEST_COUNTS.get(LEVEL, 0) == 0 %}
            --
        {% else %}
            {{ '%d, %d, %d' % 
                (100.0*np.sum(np.array(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]) <= 1) / len(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]),
                100.0*np.sum(np.array(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]) <= 10) / len(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]),
                100.0*np.sum(np.array(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]) <= 20) / len(CLASSIFIERS[clf_label].TEST_RANKINGS[LEVEL]))
            }}
        {% endif %}
        {% if LEVEL != 'PATHOGENIC' %}
            &
        {% endif %}
    {% endfor %}
    \\
{% endfor %}